From Freebase to Wikidata: The Great Migration

### 2015-10-14 paper.tex
Reviewed Sebastian's Section 4.2. Minor random changes.

\numberofauthors{5}
\author{
\alignauthor
Thomas Pellissier Tanon\titlenote{The author was an intern at Google when the majority of the work was done.}
   \affaddr{Google, San Francisco, USA}\\   
   \email{thomaspelta@gmail.com}
\alignauthor
Denny Vrandečić
   \affaddr{Google, San Francisco, USA}\\   
   \email{vrandecic@google.com}
\alignauthor
Sebastian Schaffert\\
   \affaddr{Google, Zurich, Switzerland}\\   
   \email{schaffert@google.com}
\and
\alignauthor
Thomas Steiner\\
   \affaddr{Google, Hamburg, Germany}\\   
   \email{tomac@google.com}
\alignauthor
Lydia Pintscher\\
   \affaddr{Wikimedia, Berlin, Germany}\\   
   \email{mail@pintscher.de}
}
\begin{document}

@@

\begin{abstract}
Collaborative knowledge bases that make their data freely available in a~machine-readable form
are central for the data strategy of many projects and organizations.
One such collaborative knowledge base is Freebase, initially launched by Metaweb in 2007,
who were then acquired by Google in 2010.
Another example is Wikidata, a~collaborative knowledge base developed by
Wikimedia Deutschland since 2012 and operated by the Wikimedia Foundation.
Due to the success of Wikidata, Google decided 2015 to shut down Freebase and
help the community with the transfer of Freebase's content to Wikidata.
In this paper, we describe the ongoing transfer efforts and data mapping challenges,
and provide an analysis of the effort so far.
We have created the \emph{Primary Sources Tool}
that we have released as open source to facilitate the transfer of this dataset,
but also future data migrations.
Throughout the migration, we have gained insights into both Wikidata and Freebase,
and share and discuss detailed statistics on both knowledge bases. 
\end{abstract}

@@

Moving data between two knowledge bases that do not share a~similar design
is usually a~problematic task and requires the careful mapping between their structures.
In this paper, we describe how we facilitated the migration of
Freebase\footnote{Freebase: \url{https://www.freebase.com}} content to
Wikidata.\footnote{Wikidata: \url{https://www.wikidata.org}}
That migration was no exception to this rule, and had a~number of \emph{structural} challenges.
Even more challenging was the \emph{cultural} difference between the two involved communities.
The Freebase and Wikidata communities have a~very different background,
subtly different goals and understandings of their tasks,
and different requirements regarding their data.
\todo{Expand}

@@

Freebase is an open source and
collaborative knowledge base created in 2007 by Metaweb and acquired in 2010 by Google.
It was used as the open core of the Google Knowledge Graph,
and also has many use cases outside of Google.

@@

In order to get an idea of the coverage of the two knowledge bases,
we have further created a~spatio-temporal visualization of Freebase and Wikidata 
that is shown in \autoref{fig:time-space}.
The used method is to extract the earliest date and the longitude
linked to each of the entities of the two knowledge bases
and propagate them to the connected entities which do not have such data.
To implement this propagation we set as value (date or longitude) the average of the values of the entities which distance is less than $10$ with,
as weighting, $2^{-d}$ with $d$ the distance between the two entities.
We see that the two knowledge bases follow the same basic patterns,
with their coverage mostly concentrated on Europe and North America.
There are no strong differences between them,
yet coverage of \mbox{\emph{post}-year-2000} entities seems better in Freebase.
Our theory is that this is \emph{(i)}~due to the absence of a~lot of musical data in Wikidata
that in contrast is present in Freebase,
and \emph{(ii)} due to the US population CVTs in Freebase.
The big horizontal lines in \autoref{fig:freebase} and more even in \autoref{fig:wikidata}
can most probably be traced back to places that do not have any linked date,
and in consequence inherit the foundation date of their country.
For Wikidata (\autoref{fig:wikidata}), two lines that are definitely caused by this effect
are the big line over the USA around 1780
and the one covering Russia around 850.
One possible conclusion from this visualization could be that Wikidata,
despite its smaller absolute size, appears to have
a~nearly as complete coverage of the world as Freebase.
Nonetheless, it is very challenging to find objective ways to compare the two databases
and good metrics of success for the migration.
We certainly note that for knowledge bases the amount of triples is not an adequate measure
to judge the proportion of additional good data.

>>>>

\numberofauthors{3}
\author{
Thomas Pellissier Tanon,\titlenote{The author was an intern at Google when the majority of the work was done.} Denny Vrandečić\\
   \affaddr{Google San Francisco, United States of America}\\   
   \email{\{thomaspt, vrandecic\}@google.com}
\alignauthor
Sebastian Schaffert\\
   \affaddr{Google Zurich, Switzerland}\\   
   \email{schaffert@google.com}
\alignauthor
Thomas Steiner\\
   \affaddr{Google Hamburg, Germany}\\   
   \email{tsteiner@google.com}
}

\begin{document}

@@

\begin{abstract}
Collaborative knowledge bases that make their data available in machine-readable form
as structured data are central to many companies' data strategies.
One such collaborative knowledge base is Freebase (\url{www.freebase.com}),
initially created with proprietary technologies by a~company called Metaweb,
which was then acquired by Google in 2010.
Another example is Wikidata (\url{www.wikidata.org}),
a~collaborative knowledge base developed in the open since 2012 by Wikimedia Deutschland
and operated by the Wikimedia Foundation.
Due to the success of Wikidata, in~2015, Google decided to shut down Freebase and
help the community with the transfer of its content to Wikidata through a~crowdsourcing approach.
In this paper, we describe the ongoing transfer efforts and data mapping challenges,
and provide an analysis of the process so far.
For the task, we have created a~software solution called \emph{Primary Sources Tool}
that we have released as open source to facilitate this and future data migrations.
Throughout the migration, we have gained deep insights into both Wikidata and Freebase,
and share and discuss detailed statistics on both knowledge bases. 
\end{abstract}

@@

Moving data between two knowledge bases that do not share a~similar design
is usually a~problematic task and requires the careful mapping between their structures.
In this paper, we describe how we facilitated the migration of Freebase content to Wikidata.
That migration was no exception to this rule, and had a~number of \emph{structural} challenges.
Even more challenging was the \emph{cultural} difference between the two involved communities.
The Freebase and Wikidata communities have a~very different background,
subtly different goals and understandings of their tasks,
and different requirements regarding their data.
\todo{Expand}

@@

Freebase\footnote{Freebase: \url{https://www.freebase.com/}} is an open source and
collaborative knowledge base created in 2007 by Metaweb and acquired in 2010 by Google.
It was used as the open core of the Google Knowledge Graph,
and also has many use cases outside of Google.

@@

In order to get an idea of the coverage of the two knowledge bases,
we have further created a~spatio-temporal visualization of Freebase and Wikidata 
that is shown in \autoref{fig:time-space}.
The used method is to extract the earliest date and the longitude
linked to each of the entities of the two knowledge bases
and propagate them to the connected entities which do not have such data
using a~decreasing exponential distribution \todo{Please clarify the ``propagate'' part}.
We see that the two knowledge bases follow the same basic patterns,
with their coverage mostly concentrated on Europe and North America.
There are no strong differences between them,
yet coverage of \mbox{\emph{post}-year-2000} entities seems better in Freebase.
Our theory is that this is \emph{(i)}~due to the absence of a~lot of musical data in Wikidata
that in contrast is present in Freebase,
and \emph{(ii)} due to the US population CVTs in Freebase.
The big horizontal lines in \autoref{fig:freebase} and more even in \autoref{fig:wikidata}
can most probably be traced back to places that do not have any linked date,
and in consequence inherit the foundation date of their country.
For Wikidata (\autoref{fig:wikidata}), two lines that are definitely caused by this effect
are the big line over the USA around 1780
and the one covering Russia around 850.
One possible conclusion from this visualization could be that Wikidata,
despite its smaller absolute size, appears to have
a~nearly as complete coverage of the world as Freebase.
Nonetheless, it is very challenging to find objective ways to compare the two databases
and good metrics of success for the migration.
We certainly note that for knowledge bases the amount of triples is not an adequate measure
to judge the proportion of additional good data.
